---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r}
library(readr)
library(tidyverse)
library(lubridate)
```
## Loading and preprocessing the data
```{r}
activity <- read_csv("activity.csv")
```

## What is mean total number of steps taken per day?
below is a histogram of the total number of steps taken each day
```{r}
ggplot(activity) +
    geom_histogram(mapping = aes(date, steps), stat = "identity"
                   , na.rm = TRUE)
step_per_day <- tapply(activity$steps, activity$date, FUN = sum)
step_mean <- mean(step_per_day, na.rm = TRUE)
step_median <- median(step_per_day, na.rm = TRUE)
```
the mean of the total number of steps taken per day is `r step_mean`   
the mean of the total number of steps taken per day is `r step_median`

## What is the average daily activity pattern?
```{r}
activity_avg <- activity %>% 
    group_by(interval) %>% 
    summarise(mean_step = mean(steps, na.rm = TRUE))

ggplot(activity_avg) +
    geom_line(mapping = aes(interval, mean_step))

step_max <- activity_avg[which.max(activity_avg$mean_step), 1]
```
the interval contains the maximum number of steps is `r step_max`

## Imputing missing values
```{r}
na_num <- sum(is.na(activity))
```
there are `r na_num` NA  
replace NA with the mean for the interval
```{r}
activity1 <- left_join(activity, activity_avg, "interval")
activity1[is.na(activity$steps), 1] <- 
    activity1[is.na(activity$steps), 4]

ggplot(activity1) +
    geom_histogram(mapping = aes(date, steps), stat = "identity",
                   na.rm = TRUE)
step_per_day1 <- tapply(activity1$steps, activity1$date, FUN = sum)
step_mean1 <- mean(step_per_day1, na.rm = TRUE)
step_median1 <- median(step_per_day1, na.rm = TRUE)
```
new mean of the total number of steps taken per day is `r step_mean1`  
new mean of the total number of steps taken per day is `r step_median1`  
there are little difference between the new value and the estimated value

## Are there differences in activity patterns between weekdays and weekends?
```{r}
activity1$weekday <- factor(if_else(wday(activity1$date) %in% 
                                        c(6, 7),
                                    "weekends",
                                    "weekdays"))

activity_weekday <- activity1 %>% 
    group_by(interval, weekday) %>% 
    summarise(mean_step = mean(steps, na.rm = TRUE))

ggplot(activity_weekday) +
    geom_line(mapping = aes(interval, mean_step)) +
    facet_wrap(~weekday)
```